Meta-analytic findings of the self-controlled motor learning literature: Underpowered, biased, and lacking evidential value

### 2021-09-15 docs/revision/manuscript-revised.tex
rm: duplicated code

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man,floatsintext]{apa7}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Meta-analytic findings of the self-controlled motor learning literature: Underpowered, biased, and lacking evidential value},
  pdfauthor={Brad McKay1,2, Zachary D. Yantha1, Julia Hussien1, Michael J. Carter2, \& Diane M. Ste-Marie1},
  pdflang={en-EN},
  pdfkeywords={Motor learning, retention, choice, ``OPTIMAL'\,' theory, meta-analysis, p-curve, publication bias},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\linenumbers
\usepackage{csquotes}
\usepackage{ragged2e}
\usepackage{pdflscape}
\pagewiselinenumbers
\raggedbottom
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[main=english]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{Meta-analytic findings of the self-controlled motor learning literature: Underpowered, biased, and lacking evidential value}
\author{Brad McKay\textsuperscript{1,2}, Zachary D. Yantha\textsuperscript{1}, Julia Hussien\textsuperscript{1}, Michael J. Carter\textsuperscript{2}, \& Diane M. Ste-Marie\textsuperscript{1}}
\date{}

@@

Asking learners to control any aspect of their practice environment has come to be known as self-controlled practice in the motor learning literature (Sanli, Patterson, Bray, \& Lee, 2013; Gabriele Wulf \& Lewthwaite, 2016). The first published experiments to test self-controlled learning asked learners to control their augmented feedback schedule (Janelle, Barba, Frehlich, Tennant, \& Cauraugh, 1997; Janelle, Kim, \& Singer, 1995). For example, in an experiment by Janelle, Barba, Frehlich, Tennant, and Cauraugh (1997), participants practiced throwing tennis balls at a target with their non-dominant hand. The practice period occurred over two separate days. Participants were assigned to one of four experimental groups (\emph{n} = 12): Self-controlled knowledge of performance, yoked-to-self-control, summary knowledge of performance after every five trials, and a knowledge of results only control group. The self-controlled group could request knowledge of performance whenever they wanted it, while each yoked group participant was matched with a self-control group counterpart and received knowledge of performance on the same schedule. The experimenter evaluated the participants' throws, identified the most critical error in their throwing form, and provided knowledge of performance via video feedback, along with directing attention to the error and giving prescriptive feedback. During a delayed-retention test, the accuracy, form, and speed of the throw were assessed. The results indicated that the self-control group threw more accurately and with better form than all other groups on the retention test. The self-control and yoked groups did not significantly differ in throwing speed, but the control group threw faster than the self-control group on the second retention block. The results were interpreted as evidence that the participants provided with choice were able to process information more efficiently than their counterparts who received a fixed schedule of feedback.

@@

Figure \ref{fig:fig1} shows that the number of experiments comparing self-controlled groups to yoked groups has been increasing since the original experiments by Janelle and his colleagues (1997; 1995). Researchers have experimented with giving learners control over a variety of variables in the practice environment. A qualitative assessment of the literature suggests that self-control is generally beneficial regardless of choice-type (Gabriele Wulf \& Lewthwaite, 2016). For example, self-control has been effective when participants have been provided choice over what can be considered instructionally-relevant variables, such as: knowledge of results (Patterson \& Carter, 2010), knowledge of performance (Lim et al., 2015), concurrent feedback (Huet, Camachon, Fernandez, Jacobs, \& Montagne, 2009), use of an assistive device (G. Wulf, Clauss, Shea, \& Whitacre, 2001), observation of a skilled model (Lemos, Wulf, Lewthwaite, \& Chiviacowsky, 2017), practice schedule (Will F. W. Wu \& Magill, 2011), practice volume (Lessa \& Chiviacowsky, 2015), and task difficulty (Leiker et al., 2016). Additionally, self-controlled benefits have also been found for instructionally-irrelevant variables, such as: the colour of various objects in the practice environment (Gabriele Wulf et al., 2018), other decorative choices (Iwatsuki, Navalta, \& Wulf, 2019), and the choice of what to do after the retention test is complete (Lewthwaite, Chiviacowsky, Drews, \& Wulf, 2015).

Despite the widespread optimism that self-controlled practice is useful for enhancing motor learning, researchers continue to debate the underlying mechanisms responsible for the effect (M. J. Carter \& Ste-Marie, 2017b; Gabriele Wulf et al., 2018). Beginning with Janelle, Kim, and Singer (1995), both motivational and information processing mechanisms were proposed as possible explanations for self-control benefits. Researchers have since supported these two mechanisms and, from a motivational perspective, have posited that self-control enhances confidence (Chiviacowsky, Wulf, \& Lewthwaite, 2012; Janelle, Kim, \& Singer, 1995; Gabriele Wulf \& Lewthwaite, 2016) and satisfies the basic psychological need for autonomy (Sanli, Patterson, Bray, \& Lee, 2013; Gabriele Wulf \& Lewthwaite, 2016), motivating motor performance and learning enhancement. Most self-controlled learning experiments, however, have involved participants making choices over potentially informative variables, which could act as a confounding variable. Citing this potential motivational/informational confound, Lewthwaite, Chiviacowsky, Drews, and Wulf (2015) experimented with providing instructionally-irrelevant choices, such as the colour of the golf balls to putt, the painting to hang on the wall, and what to do following the retention test. Lewthwaite and her colleagues reasoned that information processing explanations could not account for benefits due to these incidental choices, and instead motivational factors would be more likely. Consistent with the motivational hypothesis, participants exhibited significantly greater motor learning on a golf putting task (Experiment 1) and on a balance task (Experiment 2). Subsequently, several experiments have reported benefits with instructionally-irrelevant choices (Abdollahipour, Palomo Nieto, Psotta, \& Wulf, 2017; Chua, Wulf, \& Lewthwaite, 2018; Halperin, Chapman, Martin, Lewthwaite, \& Wulf, 2017; Iwatsuki, Navalta, \& Wulf, 2019; Gabriele Wulf, Chiviacowsky, \& Cardozo, 2014; Gabriele Wulf et al., 2018), further reinforcing this motivational perspective.

A contrasting line of research has been reported by Carter and his colleagues (2014; 2017a, 2017b) in which informational factors, the second dominant perspective, are given more weight as an explanatory variable. In one experiment by M. J. Carter, Carlsen, and Ste-Marie (2014), self-control participants were provided with choice over receiving knowledge of results, but divided into three experimental groups; those who could make their knowledge of results decision: before the trial, after the trial, or both (they would decide before, but could change their mind following the trial). Timing of the choice significantly attenuated the self-control benefit. While the self-after and self-both groups exhibited learning advantages relative to their yoked counterparts, the self-before group displayed no such advantage. The argument proffered by the researchers was that there was more informational value to be gained from knowledge of results requested after a trial than when it had to be requested before the outcome of the trial occurred (also see Chiviacowsky \& Wulf, 2005).

@@

Beyond this interest in the possible theoretical mechanisms, a more important question addressed was whether there is in fact evidential value for the self-controlled learning benefit. This is of relevance because the current consensus in the field is that self-controlled practice is generally more effective than yoked practice (for reviews see Sanli, Patterson, Bray, \& Lee, 2013; Ste-Marie, Carter, \& Yantha, 2019; Gabriele Wulf \& Lewthwaite, 2016). Reflecting this confidence in its benefits for motor learning, researchers have recommended adoption of self-control protocols in varied settings, such as medical training (Brydges, Carnahan, Safir, \& Dubrowski, 2009; Jowett, LeBlanc, Xeroulis, MacRae, \& Dubrowski, 2007; Gabriele Wulf, Shea, \& Lewthwaite, 2010), physiotherapy (Hemayattalab, Arabameri, Pourazar, Ardakani, \& Kashefi, 2013; Gabriele Wulf, 2007), music pedagogy (Gabriele Wulf \& Mornell, 2008), strength and conditioning (Halperin, Wulf, Vigotsky, Schoenfeld, \& Behm, 2018), and sports training (Janelle, Kim, \& Singer, 1995; Sigrist, Rauter, Riener, \& Wolf, 2013).

Problematic though is that recent, high-powered experiments with pre-registered analysis plans have failed to observe motor learning or performance benefits with self-control protocols (Grand, Daou, Lohse, \& Miller, 2017; McKay \& Ste-Marie, 2020; St. Germain et al., 2021; Yantha, 2019). Against the backdrop of the so-called replication crisis in psychology (Open Science Collaboration, 2015), there is reason for pause when evaluating the ostensible benefits of self-controlled learning. Further, Lohse, Buchanan, and Miller (2016) have raised concerns about publication bias, uncorrected multiple comparisons, \emph{p}-hacking, and other selection effects in the motor learning literature. Therefore, to address the impact of selection effects on estimates of the self-controlled learning effect, a weight function model (E. C. Carter, Schönbrodt, Gervais, \& Hilgard, 2019; Hedges \& Vevea, 1996; McShane, Böckenholt, \& Hansen, 2016; Vevea \& Hedges, 1995; Vevea \& Woods, 2005) with a one-tailed \emph{p}-value cutpoint of .025 was fit to the dataset of effects to provide a pre-registered adjusted estimate of the overall self-controlled learning effect. Even the adjusted estimate is biased if the data generating processes are biased in ways not captured by the assumptions of the model, so further sensitivity analyses were conducted to estimate the average effect of self-control after correcting for selection effects (E. C. Carter, Schönbrodt, Gervais, \& Hilgard, 2019; Vevea \& Woods, 2005). In parallel, in an effort to investigate the presence of evidential value in the literature, significant results were subjected to a \emph{p}-curve analysis (Simonsohn, Nelson, \& Simmons, 2014a; Simonsohn, Simmons, \& Nelson, 2015). The \emph{p}-curve analysis focuses exclusively on significant results and therefore is not affected by publication bias.

@@

The search commenced at PubMed and Google Scholar with the following query: *self-control** OR *self-regulat** OR *self-direct** OR *learner-control** OR *learner-regulat** OR *learner-direct** OR *subject-control** OR *subject-regulat** OR *subject-direct** OR *performer-control** OR *performer-regulat** OR *performer-direct** AND *motor learning**. The query retrieved 9014 hits on PubMed and 98,600 hits on Google Scholar. Each researcher excluded hits based on title alone or title and abstract when necessary, and quit searching the databases at self-selected intervals following extended periods of excluding 100\% of search results. Following an initial run of searching databases, each researcher employed their own search strategies, including reviewing the reference sections of reviews and included articles, consulting the ``OPTIMAL'' theory website\footnote{The webpage link that was consulted (\url{https://optimalmotorlearning.com/index.php/did-you-know-that/}) is no longer available. A new webpage devoted to ``OPTIMAL" theory can be accessed using the following link: \url{https://gwulf.faculty.unlv.edu/optimal-motor-learning/}.}, and searching the ProQuest Thesis database.

@@

The rationale for selecting the approach we did was based on five considerations. First, our interest was in motor learning as reflected by an enhanced capability to perform a skill. Motor learning studies often report multiple error measures, but they are not equally coupled with performance outcome. Constant error, for example, was not included on the priority list because it is possible to have zero constant error while performing terribly overall. Therefore, we chose to prioritize measures that could be considered to be tightly coupled with performance, like absolute error, RMSE, and absolute constant error. If these measures were not used, measures that are only correlated with performance, such as variable error, movement time, and movement form, were selected. We reasoned this selection strategy would focus the analysis on measures related to improved skill while de-emphasizing other effects. Second, we reasoned that averaging across dependent measures could introduce additional heterogeneity to the analysis by including potentially disparate dependent measures. The third, fourth, and fifth considerations all relate to avoiding bias but differ with regard to the source of the bias and the alternate method that would include such bias. Thus, the third consideration was that imposing a priority list was thought to better avoid biases that could emerge from selecting the most focal measure in a given study, because an unknowable percentage of studies may have defined the focal measure based on the strength of the findings. Fourth, we reasoned that some measures may only get reported if they support the predicted benefit of self-control. Scammacca, Roberts, and Stuebing (2014) reported that effect size estimates were inflated when random dependent measures were selected in a meta-analysis case study, perhaps reflecting a selective reporting bias. Averaging across all reported measures--a fair alternative to our approach--could conceivably pick up some of this reporting bias. Fifth, we ignored lower priority measures with data when higher priority measures lacked data because we reasoned there could be a systematic reason for this pattern: preference for reporting data associated with positive effects. Indeed, there were articles where the only measure reported with sufficient data to calculate an effect size was also the only measure with a significant result (e.g., Gabriele Wulf, Raupach, \& Pfeiffer, 2005).

@@

Selection bias in the motor learning literature is likely caused by filtering based on the statistical significance of results (Lohse, Buchanan, \& Miller, 2016). To assess and adjust for selection effects, the \texttt{R} package \texttt{weightr} (K. M. Coburn \& Vevea, 2017) was used to fit a Vevea-Hedges weight function model to the retention data (Vevea \& Hedges, 1995). The weight-function model estimates the true average effect, heterogeneity, and the probability that a non-significant result \textcolor{red}{survives censorship and is available for analysis}. Selection effects are modelled by a step function that divides the effects into two bins at one-tailed \emph{p} =.025, coinciding with a two-tailed \emph{p}-value of .05. The probability of a non-significant effect \textcolor{red}{surviving censorship to appear} in the model is estimated relative to the probability of observing a study with a significant effect. The selection-adjusted model was compared to the naïve random effects model with a likelihood ratio test. Better fit from an adjusted model suggests selection bias in the literature.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man,floatsintext]{apa7}
\title{Meta-analytic findings of the self-controlled motor learning literature: Underpowered, biased, and lacking evidential value}
\author{Brad McKay\textsuperscript{1,2}, Zachary D. Yantha\textsuperscript{1}, Julia Hussien\textsuperscript{1}, Michael J. Carter\textsuperscript{2}, \& Diane M. Ste-Marie\textsuperscript{1}}
\date{}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Meta-analytic findings of the self-controlled motor learning literature: Underpowered, biased, and lacking evidential value},
  pdfauthor={Brad McKay1,2, Zachary D. Yantha1, Julia Hussien1, Michael J. Carter2, \& Diane M. Ste-Marie1},
  pdflang={en-EN},
  pdfkeywords={Motor learning, retention, choice, ``OPTIMAL'\,' theory, meta-analysis, p-curve, publication bias},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\linenumbers
\usepackage{csquotes}
\usepackage{ragged2e}
\usepackage{pdflscape}
\pagewiselinenumbers
\raggedbottom
\ifXeTeX
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[main=english]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

@@

Asking learners to control any aspect of their practice environment has come to be known as self-controlled practice in the motor learning literature (Sanli, Patterson, Bray, \& Lee, 2013; Gabriele Wulf \& Lewthwaite, 2016). The first published experiments to test self-controlled learning asked learners to control their augmented feedback schedule (Janelle, Barba, Frehlich, Tennant, \& Cauraugh, 1997; Janelle, Kim, \& Singer, 1995). For example, in an experiment by Janelle et al. (1997), participants practiced throwing tennis balls at a target with their non-dominant hand. The practice period occurred over two separate days. Participants were assigned to one of four experimental groups (\emph{n} = 12): Self-controlled knowledge of performance, yoked-to-self-control, summary knowledge of performance after every five trials, and a knowledge of results only control group. The self-controlled group could request knowledge of performance whenever they wanted it, while each yoked group participant was matched with a self-control group counterpart and received knowledge of performance on the same schedule. The experimenter evaluated the participants' throws, identified the most critical error in their throwing form, and provided knowledge of performance via video feedback, along with directing attention to the error and giving prescriptive feedback. During a delayed-retention test, the accuracy, form, and speed of the throw were assessed. The results indicated that the self-control group threw more accurately and with better form than all other groups on the retention test. The self-control and yoked groups did not significantly differ in throwing speed, but the control group threw faster than the self-control group on the second retention block. The results were interpreted as evidence that the participants provided with choice were able to process information more efficiently than their counterparts who received a fixed schedule of feedback.

@@

Figure \ref{fig:fig1} shows that the number of experiments comparing self-controlled groups to yoked groups has been increasing since the original experiments by Janelle and his colleagues (1997, 1995). Researchers have experimented with giving learners control over a variety of variables in the practice environment. A qualitative assessment of the literature suggests that self-control is generally beneficial regardless of choice-type (Gabriele Wulf \& Lewthwaite, 2016). For example, self-control has been effective when participants have been provided choice over what can be considered instructionally-relevant variables, such as: knowledge of results (Patterson \& Carter, 2010), knowledge of performance (Lim et al., 2015), concurrent feedback (Huet, Camachon, Fernandez, Jacobs, \& Montagne, 2009), use of an assistive device (G. Wulf, Clauss, Shea, \& Whitacre, 2001), observation of a skilled model (Lemos, Wulf, Lewthwaite, \& Chiviacowsky, 2017), practice schedule (Will F. W. Wu \& Magill, 2011), practice volume (Lessa \& Chiviacowsky, 2015), and task difficulty (Leiker et al., 2016). Additionally, self-controlled benefits have also been found for instructionally-irrelevant variables, such as: the colour of various objects in the practice environment (Gabriele Wulf et al., 2018), other decorative choices (Iwatsuki, Navalta, \& Wulf, 2019), and the choice of what to do after the retention test is complete (Lewthwaite, Chiviacowsky, Drews, \& Wulf, 2015).

Despite the widespread optimism that self-controlled practice is useful for enhancing motor learning, researchers continue to debate the underlying mechanisms responsible for the effect (M. J. Carter \& Ste-Marie, 2017b; Gabriele Wulf et al., 2018). Beginning with Janelle et al. (1995), both motivational and information processing mechanisms were proposed as possible explanations for self-control benefits. Researchers have since supported these two mechanisms and, from a motivational perspective, have posited that self-control enhances confidence (Chiviacowsky, Wulf, \& Lewthwaite, 2012; Janelle et al., 1995; Gabriele Wulf \& Lewthwaite, 2016) and satisfies the basic psychological need for autonomy (Sanli et al., 2013; Gabriele Wulf \& Lewthwaite, 2016), motivating motor performance and learning enhancement. Most self-controlled learning experiments, however, have involved participants making choices over potentially informative variables, which could act as a confounding variable. Citing this potential motivational/informational confound, Lewthwaite et al. (2015) experimented with providing instructionally-irrelevant choices, such as the colour of the golf balls to putt, the painting to hang on the wall, and what to do following the retention test. Lewthwaite and her colleagues reasoned that information processing explanations could not account for benefits due to these incidental choices, and instead motivational factors would be more likely. Consistent with the motivational hypothesis, participants exhibited significantly greater motor learning on a golf putting task (Experiment 1) and on a balance task (Experiment 2). Subsequently, several experiments have reported benefits with instructionally-irrelevant choices (Abdollahipour, Palomo Nieto, Psotta, \& Wulf, 2017; Chua, Wulf, \& Lewthwaite, 2018; Halperin, Chapman, Martin, Lewthwaite, \& Wulf, 2017; Iwatsuki et al., 2019; Gabriele Wulf, Chiviacowsky, \& Cardozo, 2014; Gabriele Wulf et al., 2018), further reinforcing this motivational perspective.

A contrasting line of research has been reported by Carter and his colleagues (2014; 2017a, 2017b) in which informational factors, the second dominant perspective, are given more weight as an explanatory variable. In one experiment by M. J. Carter et al. (2014), self-control participants were provided with choice over receiving knowledge of results, but divided into three experimental groups; those who could make their knowledge of results decision: before the trial, after the trial, or both (they would decide before, but could change their mind following the trial). Timing of the choice significantly attenuated the self-control benefit. While the self-after and self-both groups exhibited learning advantages relative to their yoked counterparts, the self-before group displayed no such advantage. The argument proffered by the researchers was that there was more informational value to be gained from knowledge of results requested after a trial than when it had to be requested before the outcome of the trial occurred (also see Chiviacowsky \& Wulf, 2005).

@@

Beyond this interest in the possible theoretical mechanisms, a more important question addressed was whether there is in fact evidential value for the self-controlled learning benefit. This is of relevance because the current consensus in the field is that self-controlled practice is generally more effective than yoked practice (for reviews see Sanli et al., 2013; Ste-Marie, Carter, \& Yantha, 2019; Gabriele Wulf \& Lewthwaite, 2016). Reflecting this confidence in its benefits for motor learning, researchers have recommended adoption of self-control protocols in varied settings, such as medical training (Brydges, Carnahan, Safir, \& Dubrowski, 2009; Jowett, LeBlanc, Xeroulis, MacRae, \& Dubrowski, 2007; Gabriele Wulf, Shea, \& Lewthwaite, 2010), physiotherapy (Hemayattalab, Arabameri, Pourazar, Ardakani, \& Kashefi, 2013; Gabriele Wulf, 2007), music pedagogy (Gabriele Wulf \& Mornell, 2008), strength and conditioning (Halperin, Wulf, Vigotsky, Schoenfeld, \& Behm, 2018), and sports training (Janelle et al., 1995; Sigrist, Rauter, Riener, \& Wolf, 2013).

Problematic though is that recent, high-powered experiments with pre-registered analysis plans have failed to observe motor learning or performance benefits with self-control protocols (Grand, Daou, Lohse, \& Miller, 2017; McKay \& Ste-Marie, 2020; St. Germain et al., 2021; Yantha, 2019). Against the backdrop of the so-called replication crisis in psychology (Open Science Collaboration, 2015), there is reason for pause when evaluating the ostensible benefits of self-controlled learning. Further, Lohse, Buchanan, and Miller (2016) have raised concerns about publication bias, uncorrected multiple comparisons, \emph{p}-hacking, and other selection effects in the motor learning literature. Therefore, to address the impact of selection effects on estimates of the self-controlled learning effect, a weight function model (E. C. Carter, Schönbrodt, Gervais, \& Hilgard, 2019; Hedges \& Vevea, 1996; McShane, Böckenholt, \& Hansen, 2016; Vevea \& Hedges, 1995; Vevea \& Woods, 2005) with a one-tailed \emph{p}-value cutpoint of .025 was fit to the dataset of effects to provide a pre-registered adjusted estimate of the overall self-controlled learning effect. Even the adjusted estimate is biased if the data generating processes are biased in ways not captured by the assumptions of the model, so further sensitivity analyses were conducted to estimate the average effect of self-control after correcting for selection effects (E. C. Carter et al., 2019; Vevea \& Woods, 2005). In parallel, in an effort to investigate the presence of evidential value in the literature, significant results were subjected to a \emph{p}-curve analysis (Simonsohn, Nelson, \& Simmons, 2014b; Simonsohn, Simmons, \& Nelson, 2015). The \emph{p}-curve analysis focuses exclusively on significant results and therefore is not affected by publication bias.

@@

The search commenced at PubMed and Google Scholar with the following query: *self-control** OR *self-regulat** OR *self-direct** OR *learner-control** OR *learner-regulat** OR *learner-direct** OR *subject-control** OR *subject-regulat** OR *subject-direct** OR *performer-control** OR *performer-regulat** OR *performer-direct** AND *motor learning**. The query retrieved 9014 hits on PubMed and 98,600 hits on Google Scholar. Each researcher excluded hits based on title alone or title and abstract when necessary, and quit searching the databases at self-selected intervals following extended periods of excluding 100\% of search results. Following an initial run of searching databases, each researcher employed their own search strategies, including reviewing the reference sections of reviews and included articles, consulting the ``OPTIMAL'' theory website\footnote{The webpage link that was consulted (\url{https://optimalmotorlearning.com/index.php/did-you-know-that/}) is no longer available. A new webpage devoted to ``OPTIMAL'' theory can be accessed using the following link: \url{https://gwulf.faculty.unlv.edu/optimal-motor-learning/}.}, and searching the ProQuest Thesis database.

@@

The rationale for selecting the approach we did was based on five considerations. First, our interest was in motor learning as reflected by an enhanced capability to perform a skill. Motor learning studies often report multiple error measures, but they are not equally coupled with performance outcome. Constant error, for example, was not included on the priority list because it is possible to have zero constant error while performing terribly overall. Therefore, we chose to prioritize measures that could be considered to be tightly coupled with performance, like absolute error, RMSE, and absolute constant error. If these measures were not used, measures that are only correlated with performance, such as variable error, movement time, and movement form, were selected. We reasoned this selection strategy would focus the analysis on measures related to improved skill while de-emphasizing other effects. Second, we reasoned that averaging across dependent measures could introduce additional heterogeneity to the analysis by including potentially disparate dependent measures. The third, fourth, and fifth considerations all relate to avoiding bias but differ with regard to the source of the bias and the alternate method that would include such bias. Thus, the third consideration was that imposing a priority list was thought to better avoid biases that could emerge from selecting the most focal measure in a given study, because an unknowable percentage of studies may have defined the focal measure based on the strength of the findings. Fourth, we reasoned that some measures may only get reported if they support the predicted benefit of self-control. Scammacca et al. (2014) reported that effect size estimates were inflated when random dependent measures were selected in a meta-analysis case study, perhaps reflecting a selective reporting bias. Averaging across all reported measures--a fair alternative to our approach--could conceivably pick up some of this reporting bias. Fifth, we ignored lower priority measures with data when higher priority measures lacked data because we reasoned there could be a systematic reason for this pattern: preference for reporting data associated with positive effects. Indeed, there were articles where the only measure reported with sufficient data to calculate an effect size was also the only measure with a significant result (e.g., Gabriele Wulf, Raupach, \& Pfeiffer, 2005).

@@

Selection bias in the motor learning literature is likely caused by filtering based on the statistical significance of results (Lohse et al., 2016). To assess and adjust for selection effects, the \texttt{R} package \texttt{weightr} (K. M. Coburn \& Vevea, 2017) was used to fit a Vevea-Hedges weight function model to the retention data (Vevea \& Hedges, 1995). The weight-function model estimates the true average effect, heterogeneity, and the probability that a non-significant result \textcolor{red}{survives censorship and is available for analysis}. Selection effects are modelled by a step function that divides the effects into two bins at one-tailed \emph{p} =.025, coinciding with a two-tailed \emph{p}-value of .05. The probability of a non-significant effect \textcolor{red}{surviving censorship to appear} in the model is estimated relative to the probability of observing a study with a significant effect. The selection-adjusted model was compared to the naïve random effects model with a likelihood ratio test. Better fit from an adjusted model suggests selection bias in the literature.

The adjusted estimate from the weight-function model was pre-registered as the primary estimate of the true average effect in this meta-analysis. Please note that while the weight-function model attempts to estimate the true effect of self-controlled learning after correcting for selection biases, the estimated effect cannot be considered definitive. Nevertheless, the adjusted estimate is likely less biased than the naïve random effects estimate (E. C. Carter et al., 2019; Hong \& Reed, 2021; Kvarven, Strømland, \& Johannesson, 2020; Vevea \& Hedges, 1995). The difference between the estimates can be informative about the potential impact of selection biases, with larger disparities between models suggesting greater selection effects.
